---
title: "Class 11 Hands On -- Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1: The PDB
```{r}
pdb_stats <-read.csv("Data Export Summary.csv", row.names  = 1)
pdb_stats
percent.by.method <- pdb_stats$Total/sum(pdb_stats$Total) * 100
names(percent.by.method) <- rownames(pdb_stats)
percent.by.method
protein.percent <- print(sum(pdb_stats[,"Proteins"])/ sum(pdb_stats[,"Total"]) * 100)

```

# Q1: From the website what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers. 92.76%. 

# Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB? 1138


## HIV PDB

```{r}
pdb <- read.pdb("1hsg")
print(pdb)
```


#Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein residues? 198 AA residues, 128 non-protein residues


```{r}
pdb$atom
```


## Atom Selection

```{r}
ca.inds <- atom.select(pdb, "calpha")

```
## Protein/Ligand Selection
```{r}
prot.inds <- atom.select(pdb, string = "protein")
prot.inds
lig.inds <- atom.select(pdb, string = "ligand")
lig.inds
```

```{r}
prot.pdb <- trim.pdb(pdb, prot.inds)
write.pdb(prot.pdb, file= "protein.pdb")

```

