---
title: "Class 15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# read csv
```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
dim(counts)
```

```{r}
head(metadata)
```
# check your data more robustly
```{r}
all(colnames(counts)[-1] == metadata$id)
```

# Find which columns in counts correspond to the control conditions(i.e no drug). Then we can calculate the mean count value for each gene

```{r}
control.inds <- metadata$dex == "control"
control.ids <- metadata$id[control.inds]
```

# now that we have the control data separated, calculate the mean
```{r}
control.mean <- rowSums(counts[,control.ids]) / ncol(counts[,control.ids])
names(control.mean) <- counts$ensgene
```

# do it for the treated as well
```{r}
treated.inds <- metadata$dex == "treated"
treated.ids <- metadata$id[treated.inds]
treated.mean <- rowSums(counts[,treated.ids]) / ncol(counts[,treated.ids])
names(treated.mean) <- counts$ensgene
```

# combine the mean data into one data frame
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

```{r}
colSums(meancounts)
```

# plot the drug treated vs the control count values
```{r}
plot(log(meancounts$control.mean), log(meancounts$treated.mean))
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)
to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
sum(down.ind)
```

```{r}
head(mycounts[up.ind,])
```

```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```

```{r}
mycounts.annotated <- merge(x = mycounts, y = anno, by.x = "row.names", by.y = "ensgene")
head(mycounts.annotated)
```

## DESeq2 Analysis
```{r}
library(DESeq2)
citation("DESeq2")
```
## DESeq Pipeline
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
res <- results(dds)
summary(res)
resOrdered <- res[order(res$pvalue),]
res05 <- results(dds, alpha=0.05)

```


```{r}
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)
```

```{r}
ord <- order( resSig05$padj )
#View(res01[ord,])
head(resSig05[ord,])
```


## Volcano Plots
```{r}
res$sig <- res$padj<0.05

# How many of each?
table(res$sig)
sum(is.na(res$sig))
```

```{r}
library(ggplot2)

ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```

